Hydrodynamics and rheology of adhesive vesicle doublets

### 2019-02-05 vesicle_code/docs/adhesion/adhesion.tex
worked on quiescent example

Collective hydrodynamics of vesicles is dictated by the vesicle-vesicle
interactions.  In a quiescent flow, vesicle-vesicle adhesion leads to
the formation of vesicle doublets
\cite{Ziherl2007_PRL,ZiherlSvetina2007_PNAS} or clusters
\cite{SvetinaZiherl2008_Bioelectrochemistry,FlormannAouane2017_SciReports}.
As a model for red blood cell (RBC) aggregates, an idealized model for
adhesive vesicle-vesicle interactions gives rise to vesicle
configurations similar to those observed in experiments of
fibrinogen-induced RBC
aggregates~\cite{SvetinaZiherl2008_Bioelectrochemistry,FlormannAouane2017_SciReports}.
Using the Lennard-Jones (L-J) potential for point-point interaction
between two vesicle membranes without the molecular details of adhesive
interactions between RBCs, Flormann {\em et al.}~found the contact zone
to buckle under strong adhesion~\cite{FlormannAouane2017_SciReports}.  A
direct consequence of the buckling is the sigmoidal shape of the
adhesion region, which is also observed in RBC doublets and explained in
a slightly different model~\cite{ZiherlSvetina2007_PNAS}.  It remains
unknown how adhesive interactions between vesicles might lead to
different vesicle hydrodynamics that results in different rheological
properties of a vesicle suspension. Such studies will provide useful
insight to the rheology of blood flow and how to use nano solutes to
control the hypertension by adjusting the adhesive interactions between
RBCs.

In numerical simulations of  a vesicle suspension, the adhesive
interaction between two vesicles is often ignored
\cite{Veerapaneni2009_JCP,RahimianVeerapaneniBiros2010_JCP} for
numerical convenience, mostly because a smaller time step is often
needed to resolve the dynamics of the lubrication thin film between two
membranes with adhesion than without.  However, adhesive interactions
between RBCs lead to different RBC clusters that are expected to alter
the rheological properties of the RBC
suspension~\cite{NeuMeiselman2002_BJ,
SvetinaZiherl2008_Bioelectrochemistry}. Hydrodynamic simulations of
adhesive vesicles in a quiescent flow have revealed physical insights to
the equilibrium shapes of RBC aggregates in both
experiments~\cite{FlormannAouane2017_SciReports} and
theory~\cite{ZiherlSvetina2007_PNAS}. The main goal of this work is to
investigate the hydrodynamics of adhesive vesicles in flowing conditions
common in microfluidic vesicle flow, such as a planar shear flow and an
extensional flow.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Governing Equations \label{sec:ge}}
We consider suspensions of locally inextensible vesicles in an unbounded
two-dimensional viscous fluid.  For simplicity, we assume the fluid
viscosity both inside and outside the vesicles is constant, but our
method can be easily adjusted to account for a viscosity contrast.
Individual vesicles are denoted as $\gamma_j$, and they are
parameterized in arclength as $\xx_j(s,t)$.  The union of all vesicles
is denoted by $\gamma$.  Given a background velocity $\uu_\infty$, the
governing equations are
\begin{equation*}
\begin{aligned}
  \mu \Delta \uu = \nabla p, \quad &\xx \in \RR^2,
    &&\mbox{\em conservation of momentum} \\
  \nabla \cdot \uu = 0, \quad &\xx \in \RR^2, 
    &&\mbox{\em conservation of mass} \\
  \uu \rightarrow \uu_\infty, \quad &\|\xx\| \rightarrow \infty,
    &&\mbox{\em far-field condition} \\
  \uu(\xx,t) = \dot{\xx}, \quad &\xx \in \gamma,
    &&\mbox{\em velocity continuity} \\
  \xx_s \cdot \uu_s =0, \quad &\xx \in \gamma,
    &&\mbox{\em vesicle inextensibility} \\
  \llbracket T \rrbracket \nn = \xxi, \quad &\xx \in \gamma,
    &&\mbox{\em non-zero traction jump}
\end{aligned}
\end{equation*}
where $\xxi$ is the traction jump that depends on a bending, stretching,
and adhesion.  The bending term is $\BB \xx = -\kappa_b \xx_{ssss}$,
where $\kappa_b$ is the bending modulus, and this term comes from the
Helfrich energy.  For the numerical examples, we set $\kappa_b = 1$ and
use a Hamaker constants ranging from $0.1$ to $1$.  The stretching term
is $\TT \sigma  = (\sigma \xx_s)_s$, where $\sigma$ is the tension.  The
resistance to bending and stretching are standard assumptions of
vesicles.  In this work, we include an adhesive term, $\AA \xx$, that we
now describe.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Adhesion Model}
The adhesion model of Sukumaran and Seifert~\cite{suk-sei2001} considers
adhesion between a single three-dimensional vesicle and a solid wall at
$z=0$.  The adhesion potential in their model is of the functional form
in equation~\eqref{eq:adhesion_potential} with $(m,n) = (4,2)$:
\begin{align}
\label{eq:adhesion_potential}
  \phi(z) = H \left[ 
    \left(\frac{\delta}{z}\right)^m - \frac{m}{n} \left(\frac{\delta}{z}\right)^n \right],
\end{align}
where $z$ is the distance between the solid wall and the vesicle,
$H$ is the Hamaker constant, and $\delta$ is the adhesion length
scale.  We generalize this model to an unbounded suspension of vesicles.
We assume that the adhesion force applies between all pairs of points on
different vesicles, so we are interested in Hamaker constants that are
appropriate for lipid-lipid interactions. In
Appendix~\ref{sec:appendixA}, the adhesive force at a point $\xx$ on
vesicle $j$ is shown to be
\begin{align}
  \AA\xx:=-H m \delta^{n}\sum_{\substack{k=1 \\ k \neq j}}^M 
  \int_{\gamma_k} \frac{\xx - \yy}{\|\xx - \yy\|^{m+2}} 
  \left(\delta^{m-n} - \|\xx - \yy\|^{m-n} \right) ds_\yy.
  \label{eqn:adhesionForce}
\end{align}
The adhesion model is illustrated in Figure~\ref{fig:adhesionModel}.

\begin{figure}[htp]
  \begin{minipage}{0.31\textwidth}
    \centering
    \includegraphics[height=5cm,trim={0cm 0cm 2cm 0cm},clip]{figs/adhesionPotential.pdf}
  \end{minipage}
  \hfill
  \begin{minipage}{0.31\textwidth}
    \centering
    \includegraphics[height=5cm]{figs/configCartoon.pdf}
  \end{minipage}
  \hfill
  \begin{minipage}{0.31\textwidth}
    \centering
    \includegraphics[height=5cm,trim={3cm 0cm 2cm 0cm},clip]{figs/Adhesion_Force.png}
  \end{minipage}
  \caption{\label{fig:adhesionModel} {\em Left}: The adhesion potential
  $\phi(z)$ in Equation~\eqref{eq:adhesion_potential}.  Distances less
  than $\delta$ are repulsive and distances greater than $\delta$ are
  attractive.  {\em Center}: Each point on a vesicle is repelled or
  attracted by all points on the other vesicles.  {\em Left}: The
  resulting adhesive force is found by integrating over all points on
  all other vesicles (equation~\eqref{eqn:adhesionForce}).  The color is
  the dot product of the adhesion force with the vector joining the
  vesicles.  Therefore, the green and blue regions are repelled by the
  other vesicle while the yellow regions are attracted to the other
  vesicle.}
\end{figure}

@@

 Large value of $m$ corresponds to a sharp increase in the repulsion force as two objects are within the separation distance $\delta$.
 This poses great challenges for simulations as the numerical time-marching can be very stiff (i.e., very small time step) for large $m$. 
 The adaptive time-stepping BIE scheme makes it possible to simulate vesicle adhesive dynamics with specified numerical precision for reasonable computation time.
 We explore several combinations of $(m,n)$ in the simulations 
 of two vesicles forming a doublet in a quiescent flow.
 We found very little difference in both the dynamic evolution and equilibrium configuration between $(m,n)=(8,6)$ and $(m,n)=(4,2)$.

 %
%
% Thus we argue that a reasonable choice for $(m,n)$ is
%between $(m,n) = (4,2)$ and $(m,n) = (12,6)$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Integral Equation Formulation}
Because the fluid equations are linear and elliptic, an integral
equation formulation is possible.  This has several advantages,
including that only the vesicle interfaces require discretization.  We
start by defining the Stokes single-layer potential
\begin{align}
  \SS[\ff](\xx) &= \frac{1}{4\pi\mu} \int_\gamma \left(
    -\log \rho + \frac{\rr \otimes \rr}{\rho^2} \right) 
    \ff(\yy) ds_\yy, \quad \xx \in \RR^2,
  \label{eqn:SLP}
\end{align}
where $\rr = \xx - \yy$ and $\rho = \|\rr\|$.  The traction jump of the
velocity $\uu(\xx) = \SS[\ff](\xx)$ is the density function $\ff$, so
the fluid velocity is $\xx \in \RR^2$ is
\begin{align*}
  \uu(\xx) = \uu_{\infty}(\xx) + \SS[\xxi](\xx).
\end{align*}
The traction jump is given by three terms: the bending, tension, and
adhesion.  Therefore
\begin{align*}
  \xxi = -\kappa_b \xx_{ssss} + (\sigma \xx_s)_s + \AA \xx.
\end{align*}
Applying the no-slip boundary condition on each vesicle, the governing
equations are
\begin{align*}
  &\dot{\xx} = \uu_{\infty}(\xx) + \SS[\xxi](\xx), \\
  &\xx_s \cdot \uu_s = 0, \\
  &\xxi = -\kappa_b \xx_{ssss} + (\sigma \xx_s)_s + \AA\xx.
\end{align*}
Defining the bending operator as $\BB[\xx](\ff) = -\kappa_b \ff_{ssss}$,
and the tension operator $\TT[\xx](\sigma) = (\sigma \xx_s)_s$, the
no-slip boundary condition can be written as
\begin{align*}
  \pderiv{\xx}{t} = \SS \BB \xx + \SS \TT \sigma + \SS \AA \xx.
\end{align*}
We apply the IMEX-Euler method 
\begin{align*}
  \frac{\xx^{N+1} - \xx^N}{\Delta t} = \SS^N \BB^N \xx^{N+1} + 
  \SS^N \TT^N \sigma^{N+1} + \SS^N \AA^N \xx^N,
\end{align*}
that treats both the bending and tension terms
semi-implicitly~\cite{quaife2016adaptive}. Note that the adhesive term is
treated explicitly.  Therefore, a single time step requires solving
\begin{align}
  \xx^{N+1} - \Delta t \SS^N \BB^N \xx^{N+1} - 
    \Delta t \SS^N \TT^N \sigma^{N+1} = \xx^N + 
    \Delta t \SS^N \AA^N \xx^N,
  \label{eqn:IMEXEuler}
\end{align}
along with the inextensibility constraint that is discretized as
\begin{align*}
  \xx_s^{N} \cdot \xx_{s}^{N+1} = 1.
\end{align*}
We discretize the vesicles at a set of collocation points, compute the
bending and tension terms with Fourier differentiation, and apply
Alpert quadrature to the logarithmically singular single-layer potential
$\SS$~\cite{alp1999}.  The source and target points of the adhesion
force never coincide since they are always on different vesicles, so the
adhesion force~\eqref{eqn:adhesionForce} is computed with the spectrally
accurate trapezoid rule~\cite{tre-wei2014}.

The dynamics of a doublet undergoes many different time scales over time
horizons that are sufficiently large to characterize the formation of a
doublet and its rheological properties.  Therefore, time adaptivity is
crucial so that a user-specified tolerance is achieved with using a
guess-and-check procedure to find an appropriately small fixed time step
size.  To control the error and achieve higher-order accuracy in time,
we apply a time adaptive spectral deferred correction
method~\cite{quaife2016adaptive}.  The SDC method is able to achieve
second-order accuracy by iteratively applying the first-order
method~\eqref{eqn:IMEXEuler}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The Shear and Normal Stresses}
Consider a single vesicle $\gamma$ and a target point $\xx$ in the fluid
bulk.  In addition to the velocity $\uu$, we are also interested in
computing the shear stress and normal stresses of each vesicle.
Therefore, we must compute the total stress tensor
\begin{align*}
  T = - p I + \nabla \uu + \nabla \uu^T
\end{align*}
of the single-layer potential~\eqref{eqn:SLP}.  Given a traction jump
$\xxi$ across a vesicle $\gamma$, the velocity, pressure, and stress
tensor at a target point $\xx$ are in the bulk
\begin{align*}
  \uu(\xx) &= \SS_\uu[\xxi](\xx) = \frac{1}{4\pi\mu}\int_{\gamma} \left( 
    -\log \rho + \frac{\rr \otimes \rr}{\rho^2} \right) 
    \xxi ds_\yy, \\
    p(\xx) &= \SS_p[\xxi](\xx) = \frac{1}{2\pi} \int_{\gamma} 
    \frac{\rr \cdot \xxi}{\rho^2} ds_\yy, \\
    T(\xx) &= \SS_T[\xxi](\xx) = -\frac{1}{\pi}\int_{\gamma}
      \frac{\rr \cdot \xxi}{\rho^2} 
      \frac{\rr \otimes \rr}{\rho^2} ds_\yy.
\end{align*}
We require the stress tensor on the vesicle, and the layer potential for
$\SS_T$ has a jump as $\xx$ approaches $\gamma$.  Accounting for the
jump, the stress tensor at $\xx \in \gamma$ is
\begin{align*}
  T(\xx) = \frac{1}{2}(\nn \otimes \xxi) + \frac{1}{2} \left(
    \tt \otimes \left[
    \begin{array}{cc}
      2 t_x t_y & t_y^2 - t_x^2 \\
      t_y^2 - t_x^2 & -2 t_x t_y
    \end{array}
    \right] \xxi \right) + 
  \SS_T[\xxi](\xx).
\end{align*}
With the stress tensor computed for $\xx \in \gamma$, we can compute the
shear stress and normal stress
\begin{align*}
  \tau^\tt = (T \nn) \cdot \tt, \qquad
  \tau^\nn = (T \nn) \cdot \nn.
\end{align*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Adhesion of two vesicles in a quiescent flow} 
\label{sec:qflow}

When two vesicles move towards (or away from) each other under a
constant force $F$ without any external fluid flow, the height $(h)$ of
the thin liquid film between two vesicles follows the draining
dynamics~\cite{RamachandranLeal2010_PoF}
\begin{align}
\label{eq:dhdt}
\frac{d h}{dt} \sim \frac{K_a^{2/3} F^{1/3}}{\mu R_0^{10/3}} h^3,
\end{align}
where $K_a$ is the area expansion modulus, $\mu$ is the viscosity of
exterior fluid, $R_0$ is the radius of the undeformed vesicle.  For the
case of a constant forcing $F$ (independent of separation distance $h$
and time $t$), equation~\eqref{eq:dhdt} can be easily integrated to give
a relation between film thickness $h$ and time $t$:
\begin{align*}
t \sim \frac{\mu R_0^{10/3}}{K_a^{2/3} F^{1/3}}\frac{1}{2 h^2} \bigg|^{h(t)}_{h(0)},
\end{align*}
where $h(0)$ is the vesicle separation at $t=0$.
We note that $F< (>) 0$ for an attractive  (repulsive) interaction
between two vesicles,  and consequently $h(t)$ decreases (increases)
from  the initial separation distance $h(0)$.  When the force is
attractive, $h$ decreases monotonically and $t$ in the above equation is
the draining time that diverges as $h\rightarrow 0$.
%Similarly $F>0$ for a repulsive force and consequently $h(t) > h(0)$.

>>>>

Collective hydrodynamics of vesicles is dictated by the vesicle-vesicle
interactions.  In a quiescent flow, vesicle-vesicle adhesion leads to
the formation of vesicle doublets~\cite{Ziherl2007_PRL,
ZiherlSvetina2007_PNAS} or clusters
\cite{SvetinaZiherl2008_Bioelectrochemistry,
FlormannAouane2017_SciReports}.  As a model for red blood cell (RBC)
aggregates, an idealized model for adhesive vesicle-vesicle interactions
gives rise to vesicle configurations similar to those observed in
experiments of fibrinogen-induced RBC
aggregates~\cite{SvetinaZiherl2008_Bioelectrochemistry,
FlormannAouane2017_SciReports}.  Using the Lennard-Jones (L-J) potential
for point-point interaction between two vesicle membranes without the
molecular details of adhesive interactions between RBCs, Flormann {\em
et al.}~found the contact zone to buckle under strong
adhesion~\cite{FlormannAouane2017_SciReports}.  A direct consequence of
the buckling is the sigmoidal shape of the adhesion region, which is
also observed in RBC doublets and explained in a slightly different
model~\cite{ZiherlSvetina2007_PNAS}.  It remains unknown how adhesive
interactions between vesicles might lead to different vesicle
hydrodynamics that results in different rheological properties of a
vesicle suspension. Such studies will provide useful insight to the
rheology of blood flow and how to use nano solutes to control the
hypertension by adjusting the adhesive interactions between RBCs.

In numerical simulations of a vesicle suspension, the adhesive
interaction between two vesicles is often
ignored~\cite{Veerapaneni2009_JCP,  RahimianVeerapaneniBiros2010_JCP}
for numerical convenience, mostly because a smaller time step is often
needed to resolve the dynamics of the lubrication thin film between two
membranes with adhesion than without.  However, adhesive interactions
between RBCs lead to different RBC clusters that are expected to alter
the rheological properties of the RBC
suspension~\cite{NeuMeiselman2002_BJ,
SvetinaZiherl2008_Bioelectrochemistry}. Hydrodynamic simulations of
adhesive vesicles in a quiescent flow have revealed physical insights to
the equilibrium shapes of RBC aggregates in both
experiments~\cite{FlormannAouane2017_SciReports} and
theory~\cite{ZiherlSvetina2007_PNAS}. The main goal of this work is to
investigate the hydrodynamics of adhesive vesicles in flowing conditions
common in microfluidic vesicle flow, such as a planar shear flow and an
extensional flow.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Governing Equations \label{sec:ge}}
We consider suspensions of locally inextensible vesicles in an unbounded
two-dimensional viscous fluid.  For simplicity, we assume the fluid
viscosity both inside and outside the vesicles is constant, but our
method can be easily adjusted to account for a viscosity contrast.
Individual vesicles are denoted as $\gamma_j$, and they are
parameterized in arclength as $\xx_j(s,t)$.  The union of all vesicles
is denoted by $\gamma$.  Given a background velocity $\uu_\infty$, the
governing equations are
\begin{equation*}
\begin{aligned}
  \mu \Delta \uu = \nabla p, \quad &\xx \in \RR^2,
    &&\mbox{\em conservation of momentum} \\
  \nabla \cdot \uu = 0, \quad &\xx \in \RR^2, 
    &&\mbox{\em conservation of mass} \\
  \uu \rightarrow \uu_\infty, \quad &\|\xx\| \rightarrow \infty,
    &&\mbox{\em far-field condition} \\
  \uu(\xx,t) = \dot{\xx}, \quad &\xx \in \gamma,
    &&\mbox{\em velocity continuity} \\
  \xx_s \cdot \uu_s =0, \quad &\xx \in \gamma,
    &&\mbox{\em local inextensibility} \\
  \llbracket T \rrbracket \nn = \xxi, \quad &\xx \in \gamma,
    &&\mbox{\em non-zero traction jump}
\end{aligned}
\end{equation*}
where $\xxi$ is the traction jump that is the sum of the bending,
stretching, and adhesion forces.  The bending term $\BB \xx = -\kappa_b
\xx_{ssss}$ comes from the Helfrich energy, where $\kappa_b$ is the
bending modulus and $s$ is arclength.  For the numerical examples, we
set $\kappa_b = 1$ and consider adhesion strengths ranging from $0.1$ to
$1$ times the bending modulus.  The stretching term is $\TT \sigma  =
(\sigma \xx_s)_s$, where $\sigma$ is the tension, and the tension acts
as a Lagrange multiplier to satisfy the local extensibility constraint.
The resistance to bending and stretching are standard assumptions of
vesicles.  In this work, we include an adhesive term, $\AA \xx$, that we
now describe.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Adhesion Model}
The adhesion model of Sukumaran and Seifert~\cite{suk-sei2001} considers
adhesion between a single three-dimensional vesicle and a solid wall at
$z=0$.  They use the adhesion potential of the function form
\begin{align}
\label{eq:adhesion_potential}
  \phi(z) = H \left[ 
    \left(\frac{\delta}{z}\right)^m - \frac{m}{n} \left(\frac{\delta}{z}\right)^n \right],
\end{align}
where $z$ is the distance between the solid wall and the vesicle,
$\mathcal{H}$ is the Hamaker constant, $\delta$ is the adhesion length
scale and $(m,n) = (4,2)$.  We generalize this model to an unbounded
suspension of vesicles.  We assume that the adhesion force applies
between all pairs of points on different vesicles, so we are interested
in Hamaker constants that are appropriate for lipid-lipid interactions.
In Appendix~\ref{sec:appendixA}, the adhesive force at a point $\xx$ on
vesicle $j$ is shown to be
\begin{align}
  \AA\xx:=-\mathcal{H} m \delta^{n}\sum_{\substack{k=1 \\ k \neq j}}^M 
  \int_{\gamma_k} \frac{\xx - \yy}{\|\xx - \yy\|^{m+2}} 
  \left(\delta^{m-n} - \|\xx - \yy\|^{m-n} \right) ds_\yy.
  \label{eqn:adhesionForce}
\end{align}
The adhesion model is illustrated in Figure~\ref{fig:adhesionModel}.

\begin{figure}[htp]
  \begin{minipage}{0.31\textwidth}
    \centering
    \includegraphics[height=5cm,trim={0cm 0cm 2cm 0cm},clip]{figs/adhesionPotential.pdf}
  \end{minipage}
  \hfill
  \begin{minipage}{0.31\textwidth}
    \centering
    \includegraphics[height=5cm]{figs/configCartoon.pdf}
  \end{minipage}
  \hfill
  \begin{minipage}{0.31\textwidth}
    \centering
    \includegraphics[height=5cm,trim={3cm 0cm 2cm 0cm},clip]{figs/Adhesion_Force.png}
  \end{minipage}
  \caption{\label{fig:adhesionModel} {\em Left}: The adhesion potential
  $\phi(z)$ in equation~\eqref{eq:adhesion_potential}.  Distances less
  than $\delta$ are repulsive and distances greater than $\delta$ are
  attractive.  {\em Center}: Each point on a vesicle is repelled or
  attracted by all points on the other vesicles.  {\em Right}: The
  resulting adhesive force is found by integrating over all points on
  all other vesicles (equation~\eqref{eqn:adhesionForce}).  The color is
  the dot product of the adhesion force with the vector joining the
  vesicles.  Therefore, the green and blue regions are repelled by the
  other vesicle while the yellow regions are attracted to the other
  vesicle.}
\end{figure}

@@

Large value of $m$ corresponds to a sharp increase in the repulsion
force as two objects are within the separation distance $\delta$.  This
poses great challenges for simulations as the problem becomes stiff
(i.e., requires a very small time step) for large $m$.  The adaptive
time-stepping BIE scheme makes it possible to simulate vesicle adhesive
dynamics with specified numerical precision for reasonable computation
time.  We explore several combinations of $(m,n)$ in the simulations of
two vesicles forming a doublet in a quiescent flow.  We found very
little difference in both the dynamic evolution and equilibrium
configuration between $(m,n)=(8,6)$ and $(m,n)=(4,2)$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Integral Equation Formulation}
Because the fluid equations are linear and elliptic, a boundary integral
equation formulation is possible.  This has several advantages,
including that only the vesicle interfaces require discretization.  We
start by defining the Stokes single-layer potential
\begin{align}
  \SS[\ff](\xx) &= \frac{1}{4\pi\mu} \int_\gamma \left(
    -\log \rho + \frac{\rr \otimes \rr}{\rho^2} \right) 
    \ff(\yy) ds_\yy, \quad \xx \in \RR^2,
  \label{eqn:SLP}
\end{align}
where $\rr = \xx - \yy$ and $\rho = \|\rr\|$.  Then, given a vesicle
with a traction jump $\xxi$, the fluid velocity at is~\cite{poz1992}
\begin{align*}
  \uu(\xx) = \uu_{\infty}(\xx) + \SS[\xxi](\xx), \quad \xx \in \RR^2.
\end{align*}
The traction jump is
\begin{align*}
  \xxi = -\kappa_b \xx_{ssss} + (\sigma \xx_s)_s + \AA \xx,
\end{align*}
where the three terms correspond to the forces due to bending, tension,
and adhesion, respectively.  Therefore Applying the no-slip boundary
condition on each vesicle, the governing equations are
\begin{align*}
  &\dot{\xx} = \uu_{\infty}(\xx) + \SS[\xxi](\xx), \\
  &\xx_s \cdot \uu_s = 0, \\
  &\xxi = -\kappa_b \xx_{ssss} + (\sigma \xx_s)_s + \AA\xx.
\end{align*}
Defining the bending operator as $\BB[\xx](\ff) = -\kappa_b \ff_{ssss}$,
and the tension operator $\TT[\xx](\sigma) = (\sigma \xx_s)_s$, the
no-slip boundary condition gives
\begin{align*}
  \pderiv{\xx}{t} = \SS \BB \xx + \SS \TT \sigma + \SS \AA \xx.
\end{align*}
We apply the IMEX-Euler method 
\begin{align*}
  \frac{\xx^{N+1} - \xx^N}{\Delta t} = \SS^N \BB^N \xx^{N+1} + 
  \SS^N \TT^N \sigma^{N+1} + \SS^N \AA^N \xx^N,
\end{align*}
that treats both the bending and tension terms
semi-implicitly~\cite{qua-bir2014}. Note that the adhesive term is
treated explicitly.  Therefore, a single time step requires solving
\begin{align*}
  \xx^{N+1} - \Delta t \SS^N \BB^N \xx^{N+1} - 
    \Delta t \SS^N \TT^N \sigma^{N+1} = \xx^N + 
    \Delta t \SS^N \AA^N \xx^N,
\end{align*}
along with the inextensibility constraint that is discretized as
\begin{align*}
  \xx_s^{N} \cdot \xx_{s}^{N+1} = 1.
\end{align*}
We discretize the vesicles at a set of collocation points, compute the
bending and tension terms with Fourier differentiation, and apply Alpert
quadrature~\cite{alp1999} to the logarithmically singular single-layer
potential $\SS$.  The source and target points of the adhesion force
never coincide since they are always on different vesicles, so the
adhesion force~\eqref{eqn:adhesionForce} is computed with the spectrally
accurate trapezoid rule~\cite{tre-wei2014}.

The dynamics of a doublet undergoes many different time scales over time
horizons that are sufficiently large to characterize the formation of a
doublet and its rheological properties.  Therefore, time adaptivity is
crucial so that a user-specified tolerance is achieved without using a
guess-and-check procedure to find an appropriately small fixed time step
size.  To control the error and achieve second-order accuracy in time,
we apply a time adaptive spectral deferred correction
method~\cite{quaife2016adaptive}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The Shear and Normal Stresses}
To characterize the rheological properties of the suspension of a
doublet, it is necessary to compute shear and normal stresses.  At a
point $\xx$ in the fluid bulk, The velocity $\uu(\xx)$, pressure
$p(\xx)$, and total stress tensor $T = - p I + \nabla \uu + \nabla
\uu^T$ of the single-layer potential~\eqref{eqn:SLP} are
\begin{align*}
  \uu(\xx) &= \SS_\uu[\ff](\xx) = \frac{1}{4\pi\mu}\int_{\gamma} \left( 
    -\log \rho + \frac{\rr \otimes \rr}{\rho^2} \right) 
    \ff ds_\yy, \\
    p(\xx) &= \SS_p[\ff](\xx) = \frac{1}{2\pi} \int_{\gamma} 
    \frac{\rr \cdot \ff}{\rho^2} ds_\yy, \\
    T(\xx) &= \SS_T[\ff](\xx) = -\frac{1}{\pi}\int_{\gamma}
      \frac{\rr \cdot \ff}{\rho^2} 
      \frac{\rr \otimes \rr}{\rho^2} ds_\yy.
\end{align*}
To compute the stress tensor on a vesicle, there is a jump in the layer
potential $\SS_T$, and the stress tensor is
\begin{align*}
  T(\xx) = \frac{1}{2}(\nn \otimes \ff) + \frac{1}{2} \left(
    \tt \otimes \left[
    \begin{array}{cc}
      2 t_x t_y & t_y^2 - t_x^2 \\
      t_y^2 - t_x^2 & -2 t_x t_y
    \end{array}
    \right] \ff \right) + 
  \SS_T[\ff](\xx), \quad \xx \in \gamma,
\end{align*}
where $\nn$ and $\tt$ are the vesicle's outward unit normal and tangent
vector, respectively.  With the stress tensor computed for $\xx \in
\gamma$, the shear and normal stresses are
\begin{align*}
  \tau^\tt = (T \nn) \cdot \tt, \qquad
  \tau^\nn = (T \nn) \cdot \nn.
\end{align*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Adhesion of two vesicles in a quiescent flow} 
\label{sec:qflow} 
We consider the suspension of two identical two-dimensional vesicles
suspended in a quiescent flow.  Without any external forcing such as an imposed
electric field or a fluid flow in the far-field, the long-range
attraction pulls two vesicles together until their separation distance
is close to $\delta$.  In Section~\ref{subsec:qflow_draining_times}, we
compute the expected time for the vesicles to reach an equilibrium
configuration and compare the analysis with numerical simulations.  In
Section~\ref{subsec:qflow_adhesion_parameters}, we characterize the
effects of the adhesion parameters and the vesicles' reduced area on the
bending and adhesive energies.

\subsection{Effects of the adhesion parameters on draining times}
\label{subsec:qflow_draining_times}
When two vesicles move towards (or away from) each other under a
constant force $F$ without any external fluid flow, the height $(h)$ of
the thin liquid film between two vesicles follows the draining
dynamics~\cite{RamachandranLeal2010_PoF}
\begin{align}
  \label{eq:dhdt}
  \frac{d h}{dt} \sim \frac{K_a^{2/3} F^{1/3}}{\mu R_0^{10/3}} h^3,
\end{align}
where $K_a$ is the area expansion modulus, $\mu$ is the viscosity of the
exterior fluid, and $R_0$ is the radius of the undeformed vesicle.  For
the case of a constant forcing $F$ (independent of separation distance
$h$ and time $t$), equation~\eqref{eq:dhdt} can be easily integrated to
relate the film thickness $h$ and time $t$:
\begin{align*}
  t \sim \frac{\mu R_0^{10/3}}{K_a^{2/3} F^{1/3}}
    \frac{1}{2 h^2} \bigg|^{h(t)}_{h(0)},
\end{align*}
where $h(0)$ is the vesicle separation at $t=0$.  We note that $F< (>)
0$ for an attractive  (repulsive) interaction between two vesicles,  and
consequently $h(t)$ decreases (increases) from  the initial separation
distance $h(0)$.  When the force is attractive, $h$ decreases
monotonically, and $t$ in the above equation is the draining time that
diverges as $h\rightarrow 0$.

When the force on each vesicle is a function of the separation distance
$h$ of the form (with integers $n>m\ge 2$)
\begin{align}
  \label{eq:Fh}
  F(h) \sim \mathcal{H} \frac{\delta^m}{h^{m+1}}
    \left(\left(\frac{\delta}{h}\right)^{m-n}-1\right),
\end{align}
attraction is dominant at ``large" distances ($h > \delta$) while
repulsion is dominant at ``small" distances ($h<\delta$).  Integration
of equation~\eqref{eq:dhdt} with the $F(h)$ as defined in
equation~\eqref{eq:Fh} gives the relationship between $t$ and $h$.  In
this case the relationship involves an integral of a function of
dimensionless $\bar{h} \equiv h/\delta$:
\begin{align*}
  t\frac{K_a^{2/3}}{\mu R_0^{10/3}}& = 
    \int^{h(t)}_{h(0)}-\frac{dh}{h^3 F^{1/3}} = 
    -\frac{1}{\mathcal{H}^{1/3}\delta^{5/3}}\int^{\bar{h}(t)}_{\bar{h}(0)}
      \frac{d \bar{h}}{\bar{h}^{(8-m)/3}(\bar{h}^{m-n}-1)^{1/3}},
\end{align*}
Solving for $t$,
\begin{align}
\label{eq:teq_scaling}
  t\sim \frac{\mu R_0^{10/3}}{K_a^{2/3}}\frac{1}{\mathcal{H}^{1/3}\delta^{5/3}},
\end{align}
and equation~\eqref{eq:teq_scaling} tells us that both the adhesion
strength $\mathcal{H}$ and the separation distance $\delta$ affect the
time it takes for a pair of vesicles to reach the separation distance
$\delta$ under adhesion force $F$ in equation~\eqref{eq:Fh}.  In
particular, the duration $t$ is proportional to (i) the separation
distance $\delta^{-5/3}$ (thus $t\rightarrow \infty$ as
$\delta\rightarrow 0$), and (ii) the adhesion strength
$\mathcal{H}^{-1/3}$ (as in constant forcing
case~\cite{RamachandranLeal2010_PoF}).  From the above analysis we also
expect that the scaling of $t$ with respect to $\delta$ depends on the
adhesion potential, while the scaling with respect to adhesion strength
is independent of the specific form of the potential.

\begin{figure}
\includegraphics[keepaspectratio=true,scale=0.4]{figs/Dec13a_time_scaling01.png}
\includegraphics[keepaspectratio=true,scale=0.4]{figs/Dec13a_time_scaling02.png}
\caption{\label{fig:qflow00} Scaling of the time it takes for two
  vesicles to reach equilibrium. {\em Left}: Scaling with respect to
  $\delta$ with $\mathcal{H}=0.1$.  {\em Right}: Scaling with respect to
  $\mathcal{H}$ with
  $\delta = 0.4$. Arrows indicate the time by when the equilibrium is
  reached within $1\%$.}
  \todo[inline]{Change the $H$ in the right figure to $\mathcal{H}$}
\end{figure}
